Machine Learning for Predictive Auto-Tuning with Boosted Regression Trees

### 2011-12-14 paper/draft1.tex

One strategy for addressing these challenges is to use one of a variety of
automatic methods known collectively as ``auto-tuning.'' Two major auto-tuning
approaches have emerged in the extensive literature covering the subject (see
surveys in e.g.
\citep{vuduc2001statistical, demmel2005self, vuduc2005oski, williams2008auto,
datta2008stencil, cavazos2008intelligent, li2009note, park2011evaluation}):
analytical model-driven optimization and empirical optimization
\citep{yotov2003comparison}.

@@

\cite{pinto+cox:2011gcg} recommends as a reference implementation: block\_w 8, 
block\_h = 8, n\_filter\_rows = 1, n\_output\_4s = all, spill = $False$, imul\_fast = $True$,
pad\_shared = $True$, 
use\_text1d = $True$,
maxrreg = $\infty$,
and fast\_math = $False$.
This reference implementation was chosen manually based on good performance
across a range of platforms from older-generation cards such as the 8600GT all
the way to current-generation flagship cards such as the GTX 580 and C2070.
Given that parameters were hand-chosen for the reference kernel, no claims are
made as to the optimality nor universality of this reference (indeed, different
programmers would undoubtedly arrive at different results).  We use
this kernel configuration as a reasonable indicator of typical performance
made possible by ad hoc experimentation with parameters.

@@

For the implementation features $b$, we directly used the integer and binary
values (block\_w, block\_h, etc.) that parametrized the kernels.  We did not
use platform features ($c$) in our experiments. We leave the investigation of
cross-platform predictive auto-tuning for future work.

@@

Figure~\ref{fig:fig_gflop_scatter} shows how accurate predictive auto-tuning is compared with empirical auto-tuning.
The training set $(\mathcal{X}, \mathcal{Y})$ for the regression model comprised
all of the $<(a,b,c), y>$ pairs observed during grid search and hill-climbing search.
So for each training argument configuration $a,c$ there were 148 different values of $b$ and thus 148 training points.
Figure~\ref{fig:fig_gflop_scatter} (a) shows that as more argument configurations are used for training,
the performance of predictive auto-tuning on test configurations $(a', c)$ quickly
approaches the average performance of empirical auto-tuning.
We took care to partition the train and test sets so that there were no overlapping configurations.
The key difference between the predictive and empirical auto-tuning,
is that predictive auto-tuning
typically took about 0.1 seconds per test example, whereas empirical auto-tuning took about 1-3 minutes.

>>>>

One strategy for addressing these challenges is to use one of a variety of
automatic methods known collectively as ``auto-tuning.'' Two major auto-tuning
approaches have emer-ged in the extensive literature covering the subject (see
surveys in e.g.
\citep{vuduc2001statistical, demmel2005self, vuduc2005oski, williams2008auto,
datta2008stencil, cavazos2008intelligent, li2009note, park2011evaluation}):
analytical model-driven optimization and empirical optimization
\citep{yotov2003comparison}.

@@

\cite{pinto+cox:2011gcg} recommends as a reference implementation:
block\_w = 8, 
block\_h = 8,
n\_filter\_rows = 1,
n\_output\_4s = all,
spill = $False$,
imul\_fast = $True$,
pad\_sha-red = $True$, %%% XXX: hypenation by hand
use\_text1d = $True$,
maxrreg = $\infty$,
and fast\_math = $False$.
This reference implementation was chosen manually based on good performance
across a range of platforms from older-generation cards such as the 8600GT all
the way to current-generation flagship cards such as the GTX 580 and C2070.
Given that parameters were hand-chosen for the reference kernel, no claims are
made as to the optimality nor universality of this reference (indeed, different
programmers would undoubtedly arrive at different results).  We use
this kernel configuration as a reasonable indicator of typical performance
made possible by ad hoc experimentation with parameters.

@@

For the implementation features $b$, we directly used the integer and binary
values (block\_w, block\_h, etc.) that paramet-rized the kernels.  We did not
use platform features ($c$) in our experiments. We leave the investigation of
cross-platform predictive auto-tuning for future work.

@@

Figure~\ref{fig:fig_gflop_scatter} shows how accurate predictive auto-tuning is compared with empirical auto-tuning.
The training set $(\mathcal{X}$, $\mathcal{Y})$ for the regression model comprised
all of the ($(a,b,c)$, $y$) pairs observed during grid search and hill-climbing search.
So for each training argument configuration $a,c$ there were 148 different values of $b$ and thus 148 training points.
Figure~\ref{fig:fig_gflop_scatter} (a) shows that as more argument configurations are used for training,
the performance of predictive auto-tuning on test configurations $(a', c)$ quickly
approaches the average performance of empirical auto-tuning.
We took care to partition the train and test sets so that there were no overlapping configurations.
The key difference between the predictive and empirical auto-tuning,
is that predictive auto-tuning
typically took about 0.1 seconds per test example, whereas empirical auto-tuning took about 1-3 minutes.
